---
title: "COMP2501 practice"
output: html_notebook
author: Sunnie Shen
---

```{r}
library(tidyverse)
library(dslabs)
library(gtools)
library(ggridges) 
#for"geom_density_ridges(bindwidth=...)" plotting packed density function
library(ggplot2)
```

# Lec13-Probability

### Important: Monte Carlo simulation

## 13.1.probability and random numbers

#### Monte Carlo method to estimate pi in R

uniformly sample points, use number of points fall in circle to estimate circle area, thus calculating pi

-   runif(n_iteration, min, max): uniform sampling
-   expand.grid(): creating data.frame, and setNames(c(, )) st the same time
-   or use: data.frame to create data frame directlydf

```{r}
estimate_pi <- function(n_iter, seed= 42){
  set.seed(seed)
  x<-runif(n_iter,min=-0.5,max=0.5)#sample x axes
  y<-runif(n_iter,min=-0.5,max=0.5)#sample y axes
  dist<-sqrt(x**2+y**2)
  area = mean(dist<=0.5) # within target circle
  pi <- 4*area
  pi
}
n_iters <- c(1e1, 1e2, 1e3, 1e4, 1e5)#不同数量级的取样
seeds <- seq(1:100)
df <- expand.grid(n_iters, seeds) |> setNames(c("n_iter", "seed")) #单列,扩展出两列,形成表格
df$pi_estimates <- mapply(estimate_pi, df$n_iter, df$seed)
```

visualize the estimation for different **n_iter**, **prepare for LLN**

```{r}
df |> 
  ggplot(aes(x=n_iter, y=pi_estimates, group = n_iter)) +
  geom_hline(yintercept = pi, linetype = "dashed")+
  geom_boxplot()+ 
  scale_x_log10()
```

#### RNG(random number generaator)

-   pseudo RNG (e.g. linear congruential generator): one seed =\> deterministic result

-   True RNG : depend on physical process

## 13.2.independence and conditional probability

$P(A|B) = \frac{P(AB)}{P(B)}$ $P(AB) = {P(A)}{P(B)}$

#### Permutation: getting flush

```{r}
# Simulations for pokers
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
numbers <- c(
  "Ace", "Deuce", "Three", "Four", "Five", "Six", "Seven", 
  "Eight", "Nine", "Ten", "Jack", "Queen", "King")
deck <- expand.grid(number=numbers, suit=suits)
deck <- paste(deck$number, deck$suit)
sample(deck, 5, replace=FALSE)
```

$P(A) = P(B|A) = P(C|A) = 0.25 \Rightarrow ABC, independent$

```{r}
# Three-card scenarios
three_card_combinations <- permutations(52, 3, v=deck)

heart_cards <- paste(numbers, "Hearts")
p_first_is_heart <- mean(three_card_combinations[,1] %in% heart_cards)
p_second_is_heart <- mean(three_card_combinations[,2] %in% heart_cards)
p_third_is_heart <- mean(three_card_combinations[,3] %in% heart_cards)

p_all_are_heart <- mean(
  three_card_combinations[,1] %in% heart_cards &
  three_card_combinations[,2] %in% heart_cards &
  three_card_combinations[,3] %in% heart_cards)

print(p_first_is_heart * p_second_is_heart * p_third_is_heart)
print(p_all_are_heart)

# Event A: first card is heart
# Event B: second card is heart
# Event C: third card is heart
# P(A)
three_card_combinations |>
  as.data.frame() |>
  summarize(first_is_heart = mean(V1 %in% heart_cards))

# P(B|A)
three_card_combinations |>
  as.data.frame() |>
  filter(V1 %in% heart_cards) |>
  summarize(second_is_heart = mean(V2 %in% heart_cards))

# P(C|A,B)
three_card_combinations |>
  as.data.frame() |>
  filter(V1 %in% heart_cards & V2 %in% heart_cards) |>
  summarize(second_is_heart = mean(V3 %in% heart_cards))
```

```{R}
is_flush <- function(seed = NA){
  if(!is.na(seed)){
    set.seed(seed) #for setting random seed
  }
  hand <- sample(deck, 5, replace=FALSE) ## without replacement
  suit_in_hand <- str_split(hand, "\\s", simplify = TRUE) 
  # simplify: 回归矩阵形式, "<number> <suit(type)>"
  length(unique(suit_in_hand[,2]))==1 #same suit => flush
}
set.seed(32)
mean(replicate(1e5, is_flush()))
```

#### Combination: getting a natural 21

```{r}
# Natural 21 : getting aces and facecard in the first draw
aces <- paste("Ace", suits)
facecard <- paste(c("King", "Queen", "Jack", "Ten"), suits)

# MC simulation of getting natural 21
is_21 <- function(seed=NA){
  if (!is.na(seed)){
    set.seed(seed)
  }
  hand <- sample(deck, 2, replace=FALSE)
  flag1 <- hand[[1]] %in% aces & hand[[2]] %in% facecard
  flag2 <- hand[[2]] %in% aces & hand[[1]] %in% facecard
  flag1 | flag2
}
set.seed(42)
mean(replicate(1e5, is_21())) # probablity to get natural 21
```

## 13.3.continuous probability: expectation & variance

#### empirical CDF

```{r}
# eCDF for student heights
data(heights)
height_cdf <- function(threshold){## ecdf
  heights |> summarize(value = mean(height <= threshold)) |> pull(value)
}
height_cdf(170 / 2.53)

df <- data.frame(val = seq(120, 220, 1))
df <- df |>
  mutate(val_in_inch = val / 2.53,
         ecdf = sapply(val_in_inch, height_cdf)) ## ecdf
df |> ggplot(aes(val, ecdf)) + 
  geom_line() + 
  xlab("Height (cm)") +
  ylab("empirical CDF of student height") 
```

#### CDF of Gaussian: pnorm(抽样点序列, mean, variance)

```{r}
# Plot CDF
df <- data.frame(
  val = seq(-5, 5, 0.01), #抽样点序列
  cdf = pnorm(seq(-5, 5, 0.01), 0, 1)
)
df |> ggplot(aes(val, cdf)) +
  geom_line() + 
  ylab("CDF of a standard normal\n(mean=0, sd=1)")
```

#### PDF of normal: dnorm(抽样点序列, mean, variance)

```{r}
df <- data.frame(
  val = seq(-5, 5, 0.01),#抽样点序列
  pdf = dnorm(seq(-5, 5, 0.01), 0, 1)
)
df |> ggplot(aes(val, pdf)) + 
  geom_line() + 
  ylab("PDF of a standard normal\n(mean=0, sd=1)") 
```

#### empirical PDF: histogram, ggplot(aes()) + geom_density()

```{r}
# Histogram/density estimate of heights
height_summary <- heights |>
  filter(sex == "Male") |>
  summarize(mean=mean(height), sd=sd(height))

heights |>
  filter(sex == "Male") |>
  ggplot(aes(height)) +
  geom_density(fill="blue", alpha=0.2) +
  stat_function(
    # plot a normal distribution curve using sample mean and variance
    fun=dnorm,
    args=list(mean=height_summary$mean, sd=height_summary$sd),
    color="red") +
  ylab("Density") 
```

#### Monte Carlo simulation for continuous variable: rnorm(从normal中取样的数量, mean , variance)

```{r}
sampled_heights <- data.frame(
  height = rnorm(500, mean=height_summary$mean, sd=height_summary$sd))
sampled_heights |> ggplot(aes(height)) + geom_histogram() + theme_minimal()
```

dnorm(), dt(), dchisq(), dexp()

#### expectation & variance

$E(X)=\Sigma {x_i}{P(X_i)} ;\;Var(X)=E(X2)-E(X)2;\; SD(X)=\sqrt{Var(X)}$

## 13.4.LLN, CLT

$$
E(X)\approx \overline{X};\;
SD(X)\approx \frac{1}{n}\Sigma({X_i}P(X_i))^2;\;
$$

$$
E(\overline{X})=E(X_i);\;
SD(\overline{X})=\frac{SD(X_i)}{\sqrt{n}};\;
$$

## 13.5.Case study: Bank loan

default_rate: fail to pay back

interest_rate: target variable, too small not enough to make up bad debt, too large will result in losing existing customers

```{r}
# Case study: loan
principal_per_loan <- 180000
operational_loss <- 20000
loss_per_foreclosure <- -(principal_per_loan + operational_loss)
```

### main simulator function

```{r}
sim_average_return <- function(interest_rate=0.02, default_rate=0.02, loans=1000){
  defaults <- sample( #seq indicating default or not
      c(0,1), loans, prob=c(1-default_rate, default_rate), replace = TRUE)
  balance <- #return = revenue - cost(loss)
    sum(defaults * loss_per_foreclosure) +
      sum((1 - defaults) * principal_per_loan * interest_rate)
  balance / loans 
  #average revenue(balance)
}
```

### Case1: fixed 2% interest rate

```{r}
return_sampled_0 <- data.frame(
  return=replicate(1e5, sim_average_return(0.02))
)
#...ggplot for visulization
```

### Case2: multiple interest rate comparison

-   mutate( return = **sapply**(interest_rate, sim_average_return) )

```{r}
return_sampled_pre <- data.frame(
  interest_rate = rep(seq(0.02, 0.04, 0.001),1e5) 
  #repeating this seq 1e5 times
  #rep := repeating the first argument
)
return_sampled <- return_sampled_pre |> 
  mutate(
    return = sapply(interest_rate, sim_average_return) 
  )
```

```{r}
#错误做法
return_sample_alternative <- return_sampled_pre |>
  mutate(
    return = sim_average_return(interest_rate))|> 
  #根据sim_average_return 定义, 该函数不支持向量化输入, 必须改用sapply
  group_by(interest_rate)|>
  summarise(return = mean(return))
```

#### visualization

```{r}
# packed visualization for return-interest_rate
return_sampled |>
  mutate(interest_rate=as.factor(interest_rate)) |>
  ggplot(aes(return, interest_rate)) + 
    geom_density_ridges(bandwidth=300, fill="blue", alpha=0.2) +
    geom_vline(xintercept=0., lty="dashed") +
    theme_minimal()
#eCDF for interest_rate = 2.3%
return_sampled |>
  filter(interest_rate==0.023) |>
  ggplot(aes(return)) +
    stat_ecdf(geom = "step", color = "black") +
    geom_vline(xintercept=0., lty="dashed") +
    theme_minimal() +
    xlab("return (interest rate = 2.3%)") +
    ylab("eCDF")
# loss prob(return<0) regarding to different interest rate
return_sampled |>
  group_by(interest_rate) |>
  summarize(loss_prob = mean(return < 0)) |>
  ggplot(aes(interest_rate, loss_prob)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept=0.01, lty="dashed") +
    theme_minimal()
```

### Case4: AIM CHNAGE to — increase size of loans

$$
n\; larger \Rightarrow \; SD(average\,return)\; smaller \Rightarrow \;thinner\;tail \Rightarrow \; smaller\; Prob(return<0)
$$

```{r}
# MC at interest rate = 0.03, altering size
return_sampled <- data.frame(
  loans=rep(seq(1000, 5000, 500), 1e4))

f <- function(loans){
  sim_average_return(interest_rate=0.03, default_rate=0.02, loans=loans)
}

return_sampled <- return_sampled |>
  mutate(return=sapply(loans, f))

return_sampled |>
  group_by(loans) |>
  summarize(loss_prob = mean(return < 0)) |>
  ggplot(aes(loans, loss_prob)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.01, lty = "dashed") +
  theme_minimal()
```

## 13.6.debate between Markov and Nekrosov

-   i.i.d: independent and identical distribution

-   without i.i.d., can we get convergence(LLN) and normality(CLT)?

#### Markov chain

# Lec14-Statistical inference

## 14.1.Basics

$sample\,characteristic\Rightarrow\,intrinsic\,population\,charicteristic$

-   spread: p1-p2

-   MoE(margin of error)

```{r}
#poll from a jar of beads
p<-0.4
population_size<-1e7
jar<-sample(c(1,0),population_size,prob=c(p,1-p),replace=TRUE)#creating the population
#estimate p by sampling
take_sample<-function(n){
  sample<-sample(c(1,0),n,prob=c(p,1-p),replace=TRUE)
  mean(sample)
}

result<-data.frame(
  estimate_prop=replicate(1e4,take_sample(25))
)
result|>ggplot(aes(estimate_prop))+
  geom_histogram(binwidth=0.04)

result<-data.frame(
  estimate_prop=replicate(1e4,take_sample(1000))
)
result|>ggplot(aes(estimate_prop))+
  geom_histogram(binwidth=0.001)

#choosing binwidth: n*binwidth>1
```

## 14.2.SD,SE,CI

$$
sample\,mean: E(\overline{X})=E(X_i)=p;
$$

$$
sample\,variance: Var(\overline{X}=\frac{Var(X_i)}{n}=\frac{p(1-p)}{n}
$$

$$
standard\,error\,: SE(\overline{X})=SD(\overline{X})=\sqrt{\frac{p(1-p)}{n}}
$$

$$
Margin\,of\,error: MoE(X)=2*SE(X)
$$

-   CTL: when sample_size increases, our estimation is more "accurate", we are more "confident"\
    given that: SE decreases

```{r}
sample_sizes<-seq(1,1e4,2)
result<-data.frame(
  sample_size=sample_sizes,
  SE=sqrt(p*(1-p)/sample_sizes)
)
result|>ggplot(aes(sample_size,SE))+
  geom_point()+
  geom_line()+
  scale_x_log10()
```

$$
CLT \Rightarrow \overline{X}\sim N(\mu,\sigma)
$$

$$
95\%\,confidence: [\overline{X}\pm 1.96*SE(X)]
$$

```{r}
# Confidence interval
take_sample_and_check <- function(n){
  sampled_beads<-sample(jar,n,replace=TRUE)
  X_bar<-mean(sampled_beads)
  X_SE<-sqrt(X_bar*(1-X_bar)/n) # 95
  #assume we don't know p, so use X_bar sample parameter to estimate during calculation
  between(p,X_bar-1.96*X_SE,X_bar+1.96*X_SE)
  #<bool> check if p is within CI
}
set.seed(42)
mean(replicate(1e4, take_sample_and_check(1000)))
mean(replicate(1e4, take_sample_and_check(25)))
```

## 14.3.Power

-   (not mastered parts)

```{r}
# Power analysis at different p
ps <- c(0.501, 0.51, 0.53, 0.55, 0.6, 0.7, 0.8)
sample_size <- c(10, 30, 100, 300, 1000, 3000, 1e4, 3e4, 1e5, 3e5, 1e6, 3e6)
power_analysis <- expand.grid(p=ps, n=sample_size)
power_analysis <- power_analysis |>
  mutate(se=sqrt(p * (1-p) / n)) |>
  mutate(CI_lower=p - se * 1.95, CI_upper=p + se * 1.96)
power_analysis |>
  mutate(p = factor(p)) |>
  ggplot() +
  geom_line(aes(x=n, y=CI_lower, color=p)) +
  geom_line(aes(x=n, y=CI_upper, color=p), lty="dashed") +
  geom_hline(yintercept=0.5, color="black") +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.49, 0.51)) +
  ylab("CI on estimate of p") +
  xlab("Sample size") +
  theme_minimal()
```

## 14.4.p-value

# Lec15-Bayes theorem

## 15.1.Basics

$$
P(A|B)=\frac{P(B|A)P(A)}{P(B)}\\ 
B: observation; A:target\\
P(B):evidence证据概率\\
P(A):prior先验概率,initial,belief\\
P(B|A):likelihood似然度,证据B对A的支持程度\\
P(A|B):posterior后验概率, new,belief 
$$

## 15.2.Case study: lung cancer and pulmonary nodules

\$ B: pulmonary,nodule; A:Lung,cancer\\ Aim: P(A\|B) \$

potential bias: $P(A),P(B)$ not from same population $\Leftarrow$ condition on the same population when sampling

## 15.3.Bayesian VS frequentist

-   Bayesian modeling of coin toss: observed two heads\
    **no pre-assumption on p (different from frequentist)**

    $
    P(p|2H)=\frac{P(2H|p)P(p)}{P(2H)}\,,for\,p\in [0,1]\\
    $

```{r}
# Bayesian modeling of coin toss: observed two heads
df <- data.frame(p=seq(0, 1, 0.01))
df <- df |> mutate(posterior=p**2)

step_size <- 0.01  # Approximate the area under the curve
total_area <- sum(df$posterior * step_size)
df <- df |> mutate(posterior = posterior / total_area)  # Normalize

df |> ggplot(aes(p, posterior)) +
  geom_line() +
  ylab("Probability density (posterior)") +
  theme_minimal()
```

-   10H/14 tosses, estimate 2H in the next 2 toss\
    **no pre-assumption about p, \
    assume uniform, if prior distribution is not given**

    ```{r}
    # Bayesian modeling of coin toss: observed 10 heads out of 14 tosses
    cons = length(combinations(14, 10))
    #the combinartorial number C(14,10)
    df <- data.frame(p=seq(0, 1, 0.01))#all possible value of p
    df <- df |> mutate(posterior=cons * ((1 - p)**4) * (p**10))

    step_size <- 0.01
    total_area <- sum(df$posterior * step_size)
    df <- df |> mutate(posterior = posterior / total_area)  
    #Normalize

    df |> ggplot(aes(p, posterior)) +
      geom_line() +
      ylab("Probability density (posterior)") +
      theme_minimal()
    ```

# Lec16-Regression

## 16.1.Correlation

```{r}
library(MASS)
library(tidyverse)
install.packages("HistData")
library(HistData)
# Plotting of two variables
mu <- c(2, 5)
sigma <- matrix(c(2, 0, 0, 3), nrow = 2)  # No correlation

set.seed(42)
samples <- mvrnorm(n = 1000, mu = mu, Sigma = sigma)
samples <- data.frame(x = samples[,1], y = samples[,2])
samples |> ggplot(aes(x, y)) +
  geom_point() +
  scale_x_continuous(limits = c(-3, 7)) +
  scale_y_continuous(limits = c(-1, 9)) +
  theme_minimal()


sigma_correlated <- matrix(c(2, 1.5, 1.5, 3), nrow = 2)  # With correlation

set.seed(42)
samples <- mvrnorm(n = 1000, mu = mu, Sigma = sigma_correlated)
samples <- data.frame(x = samples[,1], y = samples[,2])
samples |> ggplot(aes(x, y)) +
  geom_point() +
  scale_x_continuous(limits = c(-3, 7)) +
  scale_y_continuous(limits = c(-1, 9)) +
  theme_minimal()
```

## 16.2.Regression

$$begin{align}
correlation\,coefficient:\rho=\frac{1}{n}\Sigma^{n}_{i=1}(\frac{x_i-\mu_x}{\sigma_x})(\frac{y_i-\mu_y}{\sigma_y})\\
regression\,line: y=b+mx\\
slope\,coefficient:m=\rho\frac{\sigma_x}{\sigma_y}=\frac{\Sigma(X_i-\overline{X})(Y_i-\overline{Y})}{\Sigma(X_i-\overline{X})^2}\\
intersection:b=\mu_y-\mu_x\\
$$

-   adding regression line

    ```{r}
    x <- galton_heights$father
    y <- galton_heights$son
    n <- length(x)
    mean(scale(x) * scale(y))
    mean((scale(x) * scale(y)) * n / (n-1))
    cor(x, y)
    cor(y, x)

    # Case study: height, two ways of calculating slope / intercept
    mu_x <- mean(x)
    s_x <- sd(x)
    mu_y <- mean(y)
    s_y <- sd(y)
    rho <- cor(x, y)

    galton_heights |>
      ggplot(aes(father, son)) +
      geom_point(alpha=0.5) +
      geom_abline(
        slope=rho * s_y/s_x,
        intercept=mu_y - mu_x * (rho * s_y/s_x)) +
      theme_minimal()

    model <- lm(son ~ father, data=galton_heights)
    galton_heights |>
      ggplot(aes(father, son)) +
      geom_point(alpha=0.5) +
      geom_abline(
        slope=model$coefficients[2],
        intercept=model$coefficients[1]) +
      theme_minimal()
    ```

## 16.3.Association
