---
title: "2501report_ShenHongshan"
author: "Shen Hongshan"
date: "2025-11-20"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Predicting Mental Health Treatment Needs Based on Workplace Factors and Demographic Characteristics

### Author: Shen Hongshan (3036290936)

### Data source: [Kaggle Mental Health in Tech Survey](https://www.kaggle.com/datasets/osmi/mental-health-in-tech-survey/data)

------------------------------------------------------------------------

## **Project Overview**

### **Proposal**

> **Research Questions:** 1. Which **workplace factors** and **demographic characteristics** are most predictive of needing mental health treatment? 2. How can we build a **regression model** to identify individuals who would **benefit from mental health intervention** based on their survey responses?

**Importance:**\
Mental health issues in the workplace have deep economic and social consequences. E**arly identification of at-risk individuals** enables timely intervention, improved well-being, and productivity.

**Challenges:**\
Handling class imbalance in responses (many may not report needing treatment). - Missing data and survey noise (subjectivity in self-reporting). - Potential bias in responses.

**Existing Work:**\
Previous studies and open-source projects have shown that factors such as *work interference*, *family history*, and *employer support* strongly impact mental health needs.

**Methodology:**\
1. Data loading and exploration

2\. Data cleaning: remove irrelevant columns, handle missingness & outliers

3\. Variable standardisation

4\. Visualisation of key factors

5\. Statistical modelling: logistic regression for predictor analysis + subgroup analyses

6\. Summary of findings and implications for workplace mental health policy

------------------------------------------------------------------------

## **1. Introduction**

### 1.1 Background

Mental health issues in the workplace have deep economic and social consequences. E**arly identification of at-risk individuals** enables timely intervention, improved well-being, and productivity.

### 1.2 Data science question

-   **What are the most significant demographics and workplace features associated with employees seeking mental health treatment?**

-   **Based on these associations, what targeted interventions can employers prioritize?**

## **2. Data Loading & Cleaning**

### 2.1 Data loading

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
library(caret) 
library(broom) 
library(car)  # for VIF
library(pROC) 
library(patchwork)

# Read the dataset (make sure 'survey.csv' is in the same folder)
survey_data <- read_csv("survey.csv")

# Overview: dimensions and column names
cat("Dataset dimensions:", dim(survey_data), "\n")
cat("Column names:\n")
names(survey_data)
```

### 2.2 Handle Missing Values, drop Irrelevant biased sampled variables

-   Check the number of missing value for each variable, delete those with too many NAs: comments, state

-   delete irrelevant variable for analysis: Timestamp

-   delete biased sampled variable: Country (60% from USA, most countries which have only 1 respondents)

```{r}
# Missing value summary
data.frame(Missing_data=sort(colSums(is.na(survey_data)),decreasing=TRUE))

# Drop columns with too many NAs or irrelevant for analysis (comments, Timestamp, state)
survey_clean0 <- survey_data |>
  select(-Timestamp, -Country, -comments, -state)

cat("Columns left after removal:", names(survey_clean))
```

### 2.3 Standardize Variables

-   Convert **Gender** variable into standard level: Male/ Female/ Transgender/Non-binary/Others

-   Remove unrealistic **Age** input, only maintain age within 18\~100 years old

-   sort the levels for better visualization latter: work_interfere(4 level), leave(5 level), no_employees(6 level)

```{r}
survey_clean0 <- survey_clean0 |>
  mutate(
    # Gender standardization
    Gender = case_when(
      str_detect(tolower(Gender), "^m|^male|^man|^maile|^make|^mal$|^male-ish|^guy") ~ "Male",
      str_detect(tolower(Gender), "^f|^female|^woman|^femaile|^femake|^fem") ~ "Female",
      TRUE ~ "Transgender/Non-binary/Others"
    ),
    # Convert character columns to factors
    across(where(is.character), as.factor)
  )

# Handle age outliers
survey_clean0 <- survey_clean0 |>
  mutate(
    Age = case_when(
      Age < 18 ~ NA,    # Remove unrealistic ages
      Age > 100 ~ NA,
      TRUE ~ Age
    )
  ) |>
  filter(!is.na(Age))  # Remove rows with invalid ages

#sort the levels of variables
survey_clean0$work_interfere <- 
  factor(survey_clean$work_interfere,levels = c("Never", "Rarely", "Sometimes", "Often"))
survey_clean0$leave <- 
  factor(survey_clean$leave,levels = c("Very difficult","Somewhat difficult","Somewhat easy","Very easy","Don't know"))
survey_clean0$no_employees <-
  factor(survey_clean$no_employees,levels = c("1-5", "6-25", "26-100", "100-500", "500-1000", "More than 1000"))
```

### 2.3 Combine variables with similar meaning

-   The variables "benefits," "care_options," "wellness_program," and "seek_help" each measure a different facet of employers' support for mental health. So, they can be consolidated into a single variable, "mental_care," with a binary (Yes/No) level.

```{r}
survey_clean <- survey_clean0|>
  mutate(mental_care=if_any(c(benefits, care_options, wellness_program, seek_help), ~ . == "Yes"))

survey_clean<- survey_clean|> 
  select(-benefits, -care_options, -wellness_program, -seek_help)
```

## **3. Exploratory Data Analysis & Visualization**

### 3.1 Target Variable Check

-   The target variable **"treatment"** indicates whether an employee is currently seeking treatment for a mental health condition.

-   **49.48%** report **No** and **50.52%** report **Yes**, which underscoring the significance of mental health conditions in the workplace.

```{r}
#Distribution of treatment status
as.data.frame(table(survey_clean$treatment))
#Proportion table
as.data.frame(prop.table(table(survey_clean$treatment)))
```

### 3.2 Demographic Overview

-   sample is skewed to male by Gender, thus it is important to notice the potential bias caused by this during analysis

```{r}
demographic_summary <- survey_clean %>%
  summarise(
    n = n(),
    avg_age = mean(Age, na.rm = TRUE),
    male_pct = round(mean(Gender == "Male")*100,1),
    female_pct = round(mean(Gender == "Female")*100,1),
    treatment_rate = round(mean(treatment == "Yes")*100,1)
  )
#Sample Demographic Summary
demographic_summary
```

### 3.3 Key Visualizations

-   Treatment rate by Age, Gender, family_history, work_interference, mental_care, tech_company

```{r}
# prepare Age-treatment information for visualization
Age_summary<-survey_clean|> 
  group_by(Age)|>
  summarise(
    n_total = n(),
    n_treatment = sum(treatment == "Yes", na.rm = TRUE),
    prop_treatment = n_treatment / n_total,
    .groups = "drop"
  )|>
  filter(n_total>=5)
#
Age_visualization<-ggplot(Age_summary , aes(x = Age, y = prop_treatment, size = n_total)) +
  geom_point(alpha = 0.6) +
  geom_smooth(alpha = 0.2)+
  theme_minimal()+
  theme(legend.position = "none")

Age_visualization
```

```{r}
# proportion of answers for different features 
feature_prop <- function(data, var, var_name) {
  # Use tidy evaluation with {{var}}
  percentage_data <- data |>
    count({{var}}) |>
    mutate(Percentage = n / sum(n))
  
  # Create plot
  ggplot(percentage_data, aes(x = {{var}}, y = Percentage, fill = {{var}})) +
    geom_bar(stat = "identity") +
    scale_y_continuous(labels = scales::percent_format(),
                       limits = c(0, max(percentage_data$Percentage) * 1.15)) +
    labs(title = paste(var_name, "of Survey Respondents"),
         x = var_name, y = "Percentage") +
    theme_minimal() +
    theme(legend.position = "none")
}

# Treatment rate by feature (function) - FIXED to use tidy evaluation
treatment_prop <- function(data, feature, feature_name) {
  ggplot(data, aes(x = {{feature}}, fill = treatment)) +
    geom_bar(position = "fill") +
    labs(y = "Proportion", x = feature_name) +
    scale_y_continuous(labels = scales::percent) +
    theme_minimal()
}

#plots for each feature
feature_prop(survey_clean, Gender, "Gender")+
  treatment_prop(survey_clean, Gender, "Gender")

feature_prop(survey_clean, family_history, "Family History")+
  treatment_prop(survey_clean, family_history, "Family History")

feature_prop(survey_clean, work_interfere, "Work Interference")+
  treatment_prop(survey_clean, work_interfere, "Work Interference")

feature_prop(survey_clean, tech_company, "Tech Company")+
  treatment_prop(survey_clean, tech_company, "Tech Company")

feature_prop(survey_clean, remote_work, "remote_work")+
  treatment_prop(survey_clean, remote_work, "remote_work")

feature_prop(survey_clean, self_employed, "self_employed")+
  treatment_prop(survey_clean, self_employed, "self_employed")

#company resourses
feature_prop(survey_clean, mental_care, "Mental Health Care")+
  treatment_prop(survey_clean, mental_care, "Mental Health Care")

feature_prop(survey_clean, anonymity, "anonymity for mental health information")+
  treatment_prop(survey_clean, anonymity, "anonymity for mental health information")

feature_prop(survey_clean, leave, "availability for taking medical leave for mental treatment")+
  treatment_prop(survey_clean, leave, "availability for taking medical leave for mental treatment")

feature_prop(survey_clean, coworkers,"feel free to discuss mental issue with coworkers")+
  treatment_prop(survey_clean,coworkers,"feel free to discuss mental issue with coworkers")
```

### **3.4 Insights**

-   Age: Treatment needs show a continuous increase from age 18, peaking at around 40 years old, followed by a significant decline in later years. This pattern may reflect several factors: increased mental health awareness and openness to treatment in middle adulthood, workplace stressors reaching their peak during career-building years, or potentially changing attitudes toward mental health across generations.
-   Gender: Females show approximately 20% higher treatment rates compared to males, while non-binary and gender-diverse individuals demonstrate over 80% need for mental treatment.
-   family_history: Those who have a family history of mental illness are around 35% more likely to get treatment rather than without a family history. This aligns with clinical understanding, as family history is a well-established risk factor for many mental health disorders.
-   Work_interfere: Respondents who report that mental health concerns "often" interfere with their work show the highest treatment rates, while those reporting "never" experiencing work interference show the lowest. This suggests that work interference serves as a significant motivator for seeking treatment, highlighting the **importance of workplace mental health interventions.**
-   tech_company/ self_employed: The little difference in treatment rates between tech company employees, non-tech employees, and self-employed individuals suggests that industry type and employment structure have limited influence on treatment-seeking behavior.
-   Mental healthcare/ anonymity/ leave: Companies that provide comprehensive mental health care programs—including confidential counseling services, robust privacy protections for mental health information, and clear mental health leave policies—show significantly higher treatment rates among employees. This supports the implementation of workplace mental health programs and policies are effective strategy for increasing treatment engagement.

## **4. Modelling & Statistical Analysis**

### 4.1 Prepare Data for Modelling

```{r}
# Create binary target variable
survey_model <- survey_clean %>%
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )

# Select key predictors as per project proposal
predictors <- c(
  "Age", "Gender", "family_history",
  "work_interfere", "no_employees", "remote_work", "tech_company",
  "mental_care",
  "anonymity", "leave",
  "mental_health_consequence", "phys_health_consequence",
  "coworkers", "supervisor", "mental_health_interview",
  "mental_vs_physical", "obs_consequence"
)

model_data <- survey_model |>
  select(all_of(c("treatment_binary", predictors))) |>
  na.omit() ##删除数据缺失行

head(model_data)
```

### 4.2 Split Data: Train/Test

```{r}
set.seed(42)
train_idx <- createDataPartition(model_data$treatment_binary, p=0.7, list=FALSE)
train_data <- model_data[train_idx, ]
test_data <- model_data[-train_idx, ]

cat("Training set size:", nrow(train_data), "\n")
cat("Test set size:", nrow(test_data), "\n")
```

### 4.3 Logistic Regression Model

target variable & feature variable are both discrete, most of them with only 2\~3 levels. So, we can not use linear regression, but logistic regression to explore what are the significant features associated with our target value "treatment"

```{r}
logit_model <- glm(treatment_binary ~ ., data=train_data, family="binomial")

cat("Model summary (logistic regression):\n")
summary(logit_model)
```

#### Predictors sorted by p-value (significance)

```{r}
significant_predictors <- broom::tidy(logit_model) |>
  arrange(p.value)
significant_predictors
```

------------------------------------------------------------------------

#### Feature Importance Plot

```{r}
feature_importance <- broom::tidy(logit_model) %>%
  filter(term != "(Intercept)") %>%
  mutate(odds_ratio=exp(estimate),
         direction=ifelse(estimate>0,"Increased","Decreased")) %>%
  arrange(desc(abs(estimate)))
top_features <- feature_importance %>% head(15)

ggplot(top_features, aes(x=reorder(term, estimate), y=estimate, fill=direction)) +
  geom_col() +
  coord_flip() +
  labs(title="Top 15 Most Important Predictors of Mental Health Treatment Need",
       x="Predictor", y="Estimate (log-odds)", fill="Direction") +
  theme_minimal()
```

### 4.4 Insights

-   **Work interference** is the single strongest signal: respondents saying their mental health often interferes with work are far more likely to be in treatment. This supports screening based on self‑reported work interference.

-   **Family history** and **coworker openness** are also meaningful predictors. Family history increases risk, and feeling free to discuss with coworkers is positively associated with being in treatment, reflecting greater willingness to seek help or workplaces where issues are more openly discussed.

-   Company provisions (**mental_care TRUE, anonymity TRUE**) are positively associated with treatment.This can mean:

    -    access to care increases uptake

    -   companies experiencing higher demand put programs in place (reverse causality)

    -   confounding by unmeasured factors.

-   Male gender is associated with lower odds of being in treatment. That likely reflects **differing help‑seeking behavior**, **not necessarily lower need.**

### 5. Model Evaluation

#### ROC Curve

-   ROC plots **sensitivity (true positive rate)** vs **1 − specificity (false positive rate)** over thresholds. The curve is **well above the diagonal:** model discriminates much better than random.

-   **AUC = 0.8476716** indicates good ranking/discrimination: the model usually gives higher predicted probabilities to respondents who are actually in treatment than to those who are not.

```{r}
test_data$predicted_prob <- predict(logit_model, newdata=test_data, type="response")
test_data$predicted_class <- ifelse(test_data$predicted_prob > 0.5, 1, 0)

test_conf <- confusionMatrix(as.factor(test_data$predicted_class), as.factor(test_data$treatment_binary))
print(test_conf)

roc_curve <- roc(test_data$treatment_binary, test_data$predicted_prob)
cat("AUC on test:", auc(roc_curve), "\n")
plot(roc_curve, main="ROC Curve for Mental Health Treatment Prediction", col="blue", lwd=3)
abline(a=0, b=1, col="grey", lty=2)
```

#### Prediction (training set)

-   Strength of model:\
     **High Specificity (91.91%)**: Excellent at identifying people who truly need treatment\
     **Good Overall Accuracy (80.23%)**: Beats baseline by 16 percentage points\
     **Balanced Precision**: Similar precision for both classes (\~80%)\
     **Statistical Significance**: Results are not due to chance

-   Concerns\
     **Low Sensitivity (59.27%)**: Poor at identifying people who don't need treatment\
     **Imbalanced Performance**: Much better at detecting treatment needs than non-needs\
     **Potential Over-intervention**: 40.73% of "No treatment" people are incorrectly flagged

```{r}
train_data$predicted_prob <- predict(logit_model, type="response")
train_data$predicted_class <- ifelse(train_data$predicted_prob > 0.5, 1, 0)
train_conf <- confusionMatrix(as.factor(train_data$predicted_class), as.factor(train_data$treatment_binary))
print(train_conf)

# vusualize confusion data
conf_matrix_data <- as.data.frame(as.table(train_conf$table))
names(conf_matrix_data) <- c("Predicted", "Actual", "Freq")

# calculate relevant statistics
conf_matrix_data <- conf_matrix_data %>%
  group_by(Actual) %>%
  mutate(
    Percentage = Freq / sum(Freq) * 100,
    Label = paste0(Freq, "\n(", round(Percentage, 1), "%)")
  )

# plot heatmap ofr confusion matrix
ggplot(conf_matrix_data, aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile(color = "white", alpha = 0.8) +
  geom_text(aes(label = Label), color = "black", size = 5) +
  scale_fill_gradient(low = "#F8F9FA", high = "blue", name = "frequency") +
  labs(
    title = "heatmap ofr confusion matrix",
    subtitle = paste("Accuracy:", round(train_conf$overall["Accuracy"]*100, 1), "%"),
    x = "Actural class",
    y = "Predicted class",
    caption = paste("sensitivity:", round(train_conf$byClass["Sensitivity"]*100, 1), "% |",
                   "Specificity:", round(train_conf$byClass["Specificity"]*100, 1), "%")
  ) +
  theme_minimal() 
```

## **6. Answering data science questions**

### 6.1. How to identify potential mental issue, In other words, what features are most predictive of needing mental health treatment?

-   We identified that **family history**, **work interference**, and **workplace support systems** are among the strongest predictors of mental health treatment need. (detailed analysis seen in **section 3.4 & 4.4**)

### 6.2 What could employers do to reduce mental issue's risks?

-   A logistic regression model giving **the most predictive factors** can reasonably **identify potnetial employees at higher risk.**
-   Companies should focus on improving **support policies** for employees showing risk factors.
    -   easier access to a**pplying for medical leave due to mental issue**
    -   protect workers **anonymity regarding mental health issue**, and creating a non-judgmental environment
-   Increasing mental health care benefits, easy-access mental care treatment resources.
    -   **interesting** to notice the **different approach** that may apply to different Gender (see visualization below!)
        -   **women are more sensitive to mental care benefit (money)**: providing more money on mental health care have a greater positive effect on seeking treatment for women(nearly **30%** more) compared to men(around **10%** more).
        -   on the other hand, **men are more sensitive to giving health care access**: providing more money on mental health care have a greater positive effect on seeking treatment for men(around **15%** increase) compared to women(lower than **5%** increase).
        -   giving **direct mental care options at work place**, is the **most effective** way to encourage employees seeking treatment, for both men and women.

```{r}
# 1. Interaction Effects Visualization - benefit * Gender
gender_care_data <- survey_clean0 |>
  filter(Gender!="Transgender/Non-binary/Others")|>
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )|>
  group_by(Gender, benefits) |>
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

p_benefits<-ggplot(gender_care_data, aes(x = Gender, y = treatment_rate, fill = benefits)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × mental health benefits",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nMental Health Benefits"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")


# 2. Interaction Effects Visualization - seek_help * Gender
gender_care_data <- survey_clean0 |>
  filter(Gender!="Transgender/Non-binary/Others")|>
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )|>
  group_by(Gender, seek_help) |>
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

p_seek_help<-ggplot(gender_care_data, aes(x = Gender, y = treatment_rate, fill = seek_help)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × Access to help",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nAccess to help"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")

# 3. Interaction Effects Visualization - care_options * Gender
gender_care_data <- survey_clean0 |>
  filter(Gender!="Transgender/Non-binary/Others")|>
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )|>
  group_by(Gender, care_options) |>
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

p_care_options<-ggplot(gender_care_data, aes(x = Gender, y = treatment_rate, fill = care_options)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × mental care options",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nMental care options"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")

p_benefits + p_seek_help + p_care_options
```

## 7. Limitation & possible improvement

-   **Most significant limitation:** Subjective survey responses, possible response bias.
    -   target variable "treatment" demonstrate the actual need for mental treatment (**real mental issue posed**). However, due to subjective survey response
-   Future work could include more advanced ML (random forest, boosting), longitudinal data, and testing interventions.

------------------------------------------------------------------------

### 7.1 Simple correlation to interaction effects

```{r}
# 1. Interaction Effects Visualization - work_interfere * family_history
interaction_data <- train_data %>%
  group_by(work_interfere, family_history) %>%
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  ) %>%
  filter(!is.na(work_interfere))

ggplot(interaction_data, aes(x = work_interfere, y = treatment_rate, 
                             color = family_history, group = family_history)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            vjust = -0.8, size = 3) +
  labs(
    title = "Interaction Effect: Work Interference × Family History",
    subtitle = "Family history significantly amplifies the impact of work interference on treatment needs",
    x = "Frequency of Work Interference",
    y = "Treatment Rate",
    color = "Family History of\nMental Health"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("No" = "red", "Yes" = "blue"))
```

-   female is more sensitive to male regarding mental heath benefit

```{r}

```

```{r}
# 2. Interaction Effects Visualization - care_options * Gender
gender_care_data <- survey_clean0 |>
  filter(Gender!="Transgender/Non-binary/Others")|>
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )|>
  group_by(Gender, care_options) |>
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

ggplot(gender_care_data, aes(x = Gender, y = treatment_rate, fill = care_options)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × mental care options",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nMental care options"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")


```

```{r}
# 2. Interaction Effects Visualization - seek_help * Gender
gender_care_data <- survey_clean0 |>
  filter(Gender!="Transgender/Non-binary/Others")|>
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )|>
  group_by(Gender, seek_help) |>
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

ggplot(gender_care_data, aes(x = Gender, y = treatment_rate, fill = seek_help)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × Access to help",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nAccess to help"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")
```

**"When we dug deeper into our analysis, we discovered a crucial insight: risk factors don't operate in isolation - they amplify or buffer each other.”**

**The first chart reveals** a powerful interaction between **work interference and family history**:

-   For employees without family history, even frequent work interference leads to only 60% treatment rate

-   But for those with family history, the same level of work interference drives treatment rates above 90%

-   **Business Insight**: This tells us screening should **prioritize employees who report both family history AND work interference**

**The second chart shows** how **gender and (mental health)benefits** interact to reveal important organizational patterns:

-   Male employees are highly sensitive to benefits availability - treatment rates drop sharply without benefits

```{r correlation-heatmap, warning=FALSE, message=FALSE}
# 说明：
# - 我们把因子变量临时编码为数值（as.numeric(as.factor(.))）来计算 Spearman 相关系数。
# - 这只是用于“可视化与探索”目的：因子->数值的转换会改变变量含义，不能当作严格的相关性/因果性证据。
# - 对于类别-类别的关联，更合适的方法是 Cramer's V；对于数值-二分类可用 point-biserial 等。
#
# 需要的包：corrplot, reshape2, Hmisc（可选，用于 p 值）
if(!requireNamespace("corrplot", quietly=TRUE)) install.packages("corrplot")
if(!requireNamespace("reshape2", quietly=TRUE)) install.packages("reshape2")
if(!requireNamespace("Hmisc", quietly=TRUE)) install.packages("Hmisc")

library(corrplot); library(reshape2); library(Hmisc)

# 选取用于可视化的变量（使用在建模中挑选的 predictors）
corr_df <- model_data 

# 把因子变量临时编码为数值（仅用于计算相关矩阵）
corr_num <- corr_df |>
  mutate(across(everything(), ~ if(is.factor(.)) as.numeric(as.factor(.)) else .))

# 计算 Spearman 相关系数（对有序/非正态数据更稳健）
corr_mat <- cor(corr_num, use = "pairwise.complete.obs", method = "spearman")

corr_mat[upper.tri(corr_mat, diag = TRUE)] <- NA

# 备用：用 ggplot2 绘制完整的热图（旋转标签更友好）
melted <- reshape2::melt(corr_mat, na.rm = TRUE)
ggplot(melted, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limits = c(-0.5,0.6)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "", fill = "Spearman\nrho",
       title = "Spearman Correlation Heatmap (factors -> numeric)")
```

## 8. References

OSMI Mental Health in Tech Survey [Kaggle Dataset](https://www.kaggle.com/datasets/osmi/mental-health-in-tech-survey/data) - [R for Data Science](https://r4ds.had.co.nz/) - [Tidyverse Documentation](https://www.tidyverse.org/) - [Introduction to Statistical Learning](https://www.statlearning.com/)

### 
