---
title: "2501report_ShenHongshan"
author: "Shen Hongshan"
date: "2025-11-20"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Predicting Mental Health Treatment Needs Based on Workplace Factors and Demographic Characteristics

### Author: Shen Hongshan (3036290936)

### Data source: [Kaggle Mental Health in Tech Survey](https://www.kaggle.com/datasets/osmi/mental-health-in-tech-survey/data)

------------------------------------------------------------------------

## **Project Overview**

### **Proposal**

> **Research Questions:** 1. Which **workplace factors** and **demographic characteristics** are most predictive of needing mental health treatment? 2. How can we build a **regression model** to identify individuals who would **benefit from mental health intervention** based on their survey responses?

**Importance:**\
Mental health issues in the workplace have deep economic and social consequences. E**arly identification of at-risk individuals** enables timely intervention, improved well-being, and productivity.

**Challenges:**\
Handling class imbalance in responses (many may not report needing treatment). - Missing data and survey noise (subjectivity in self-reporting). - Potential bias in responses.

**Existing Work:**\
Previous studies and open-source projects have shown that factors such as *work interference*, *family history*, and *employer support* strongly impact mental health needs.

**Methodology:**\
1. Data loading and exploration

2\. Data cleaning: remove irrelevant columns, handle missingness & outliers

3\. Variable standardisation

4\. Visualisation of key factors

5\. Statistical modelling: logistic regression for predictor analysis + subgroup analyses

6\. Summary of findings and implications for workplace mental health policy

------------------------------------------------------------------------

## **1. Introduction**

### 1.1 Background

Mental health issues in the workplace have deep economic and social consequences. E**arly identification of at-risk individuals** enables timely intervention, improved well-being, and productivity.

### 1.2 Data science question

-   **What are the most significant demographics and workplace features associated with employees seeking mental health treatment?**

-   **Based on these associations, what targeted interventions can employers prioritize?**

## **2. Data Loading & Cleaning**

### 2.1 Data loading

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
library(caret) 
library(broom) 
library(car)  # for VIF
library(pROC) 

# Read the dataset (make sure 'survey.csv' is in the same folder)
survey_data <- read_csv("survey.csv")

# Overview: dimensions and column names
cat("Dataset dimensions:", dim(survey_data), "\n")
cat("Column names:\n")
names(survey_data)
```

### 2.2 Handle Missing Values, drop Irrelevant biased sampled variables

-   Check the number of missing value for each variable, delete those with too many NAs: comments, state

-   delete irrelevant variable for analysis: Timestamp

-   delete biased sampled variable: Country (60% from USA, mostcountries which have only 1 respondents)

```{r}
# Missing value summary
data.frame(Missing_data=sort(colSums(is.na(survey_data)),decreasing=TRUE))

# Drop columns with too many NAs or irrelevant for analysis (comments, Timestamp, state)
survey_clean <- survey_data |>
  select(-Timestamp, -Country, -comments, -state)

cat("Columns left after removal:", names(survey_clean))
```

### 2.3 Standardize Variables

-   Convert gender variable into standard level: Male/ Female/ Transgender/Non-binary/Others

-   remove unrealistic age input, only maintain age within 18\~100 years old

```{r}
survey_clean <- survey_clean |>
  mutate(
    # Gender standardization
    Gender = case_when(
      str_detect(tolower(Gender), "^m|^male|^man|^maile|^make|^mal$|^male-ish|^guy") ~ "Male",
      str_detect(tolower(Gender), "^f|^female|^woman|^femaile|^femake|^fem") ~ "Female",
      TRUE ~ "Transgender/Non-binary/Others"
    ),
    # Convert character columns to factors
    across(where(is.character), as.factor)
  )

# Handle age outliers
survey_clean <- survey_clean |>
  mutate(
    Age = case_when(
      Age < 18 ~ NA,    # Remove unrealistic ages
      Age > 100 ~ NA,
      TRUE ~ Age
    )
  ) |>
  filter(!is.na(Age))  # Remove rows with invalid ages
```

### 2.3 Combine variables with similar meaning

-   The variables "benefits," "care_options," "wellness_program," and "seek_help" each measure a different facet of employers' support for mental health. So, they can be consolidated into a single variable, "mental_care," with a binary (Yes/No) level.

```{r}
survey_clean <- survey_clean|>
  mutate(mental_care=if_any(c(benefits, care_options, wellness_program, seek_help), ~ . == "Yes"))

survey_clean<- survey_clean|> 
  select(-benefits, -care_options, -wellness_program, -seek_help)
```

## **3. Exploratory Data Analysis & Visualization**

### 3.1 Target Variable Check

-   The target variable **"treatment"** indicates whether an employee is currently seeking treatment for a mental health condition.

-   **49.48%** report **No** and **50.52%** report **Yes**, which underscoring the significance of mental health conditions in the workplace.

```{r}
#Distribution of treatment status
as.data.frame(table(survey_clean$treatment))
#Proportion table
as.data.frame(prop.table(table(survey_clean$treatment)))
```

### 3.2 Demographic Overview

-   sample is skewed to male by Gender, thus it is important to notice the potential bias caused by this during analysis

```{r}
demographic_summary <- survey_clean %>%
  summarise(
    n = n(),
    avg_age = mean(Age, na.rm = TRUE),
    male_pct = round(mean(Gender == "Male")*100,1),
    female_pct = round(mean(Gender == "Female")*100,1),
    treatment_rate = round(mean(treatment == "Yes")*100,1)
  )
#Sample Demographic Summary
demographic_summary
```

### 3.2 Key Visualisations

-   Treatment rate by Age, Gender, family history, work_interference, mental_care, tech_company

```{r, cache=TRUE}
Age_summary<-survey_clean|> 
  group_by(Age)|>
  summarise(
    n_total = n(),
    n_treatment = sum(treatment == "Yes", na.rm = TRUE),
    prop_treatment = n_treatment / n_total,
    .groups = "drop"
  )|>
  filter(n_total>=5)

Age_visualization<-ggplot(Age_summary , aes(x = Age, y = prop_treatment, size = n_total)) +
  geom_point(alpha = 0.6) +
  geom_smooth(alpha = 0.2)+
  theme_minimal()+
  theme(legend.position = "none")


Exploratory_visualization<-function(feature,feature_name){
  ggplot(survey_clean, aes(x=feature, fill=treatment)) +
  geom_bar(position="fill") +
  labs(y="Proportion",x=feature_name) +
  scale_y_continuous(labels=scales::percent) +
  theme_minimal()
}

(Age_visualization+
    Exploratory_visualization(survey_clean$Gender, "Gender")+
    Exploratory_visualization(survey_clean$family_history, "family_history"))/
  (Exploratory_visualization(survey_clean$work_interfere, "work_interfere")+
     Exploratory_visualization(survey_clean$mental_care, "mental_care")+
     Exploratory_visualization(survey_clean$tech_company, "tech_company"))
```

#### **Insight:**

*Individuals with family history of mental illness and those reporting frequent work interference are much more likely to seek treatment.*

## **4. Modelling & Statistical Analysis**

### 4.1 Prepare Data for Modelling

```{r}
# Create binary target variable
survey_model <- survey_clean %>%
  mutate(
    treatment_binary = ifelse(treatment == "Yes", 1, 0)
  )

# Select key predictors as per project proposal
predictors <- c(
  "Age", "Gender", "family_history",
  "work_interfere", "no_employees", "remote_work", "tech_company",
  "mental_care",
  "anonymity", "leave",
  "mental_health_consequence", "phys_health_consequence",
  "coworkers", "supervisor", "mental_health_interview",
  "mental_vs_physical", "obs_consequence"
)

model_data <- survey_model |>
  select(all_of(c("treatment_binary", predictors))) |>
  na.omit() ##删除数据缺失行

head(model_data)
```

### 4.2 Split Data: Train/Test

```{r}
set.seed(42)
train_idx <- createDataPartition(model_data$treatment_binary, p=0.7, list=FALSE)
train_data <- model_data[train_idx, ]
test_data <- model_data[-train_idx, ]

cat("Training set size:", nrow(train_data), "\n")
cat("Test set size:", nrow(test_data), "\n")
```

------------------------------------------------------------------------

### 4.3 Logistic Regression Model

target variable & feature variable are both discrete, most of them with only 2\~3 levels. So, we can not use linear regression, but logistic regression to explore what are the significant features associated with our target value "treatment"

```{r}
logit_model <- glm(treatment_binary ~ ., data=train_data, family="binomial")

cat("Model summary (logistic regression):\n")
summary(logit_model)
```

#### Significant predictors (p \< .05)

```{r}
significant_predictors <- broom::tidy(logit_model) |>
  arrange(p.value)
significant_predictors
```

------------------------------------------------------------------------

#### Feature Importance Plot

```{r}
feature_importance <- broom::tidy(logit_model) %>%
  filter(term != "(Intercept)") %>%
  mutate(odds_ratio=exp(estimate),
         direction=ifelse(estimate>0,"Increased","Decreased")) %>%
  arrange(desc(abs(estimate)))
top_features <- feature_importance %>% head(15)

ggplot(top_features, aes(x=reorder(term, estimate), y=estimate, fill=direction)) +
  geom_col() +
  coord_flip() +
  labs(title="Top 15 Most Important Predictors of Mental Health Treatment Need",
       x="Predictor", y="Estimate (log-odds)", fill="Direction") +
  theme_minimal()
```

### 5. Model Evaluation

#### ROC Curve

AUC=0.8476716,

```{r}
test_data$predicted_prob <- predict(logit_model, newdata=test_data, type="response")
test_data$predicted_class <- ifelse(test_data$predicted_prob > 0.5, 1, 0)

test_conf <- confusionMatrix(as.factor(test_data$predicted_class), as.factor(test_data$treatment_binary))
print(test_conf)

roc_curve <- roc(test_data$treatment_binary, test_data$predicted_prob)
cat("AUC on test:", auc(roc_curve), "\n")
plot(roc_curve, main="ROC Curve for Mental Health Treatment Prediction", col="#2c3e50", lwd=3)
abline(a=0, b=1, col="grey", lty=2)
```

#### Prediction (training set)

-   Strength of model:\
     **High Specificity (91.91%)**: Excellent at identifying people who truly need treatment\
     **Good Overall Accuracy (80.23%)**: Beats baseline by 16 percentage points\
     **Balanced Precision**: Similar precision for both classes (\~80%)\
     **Statistical Significance**: Results are not due to chance

-   Concerns\
     **Low Sensitivity (59.27%)**: Poor at identifying people who don't need treatment\
     **Imbalanced Performance**: Much better at detecting treatment needs than non-needs\
     **Potential Over-intervention**: 40.73% of "No treatment" people are incorrectly flagged

```{r}
train_data$predicted_prob <- predict(logit_model, type="response")
train_data$predicted_class <- ifelse(train_data$predicted_prob > 0.5, 1, 0)
train_conf <- confusionMatrix(as.factor(train_data$predicted_class), as.factor(train_data$treatment_binary))
print(train_conf)

# vusualize confusion data
conf_matrix_data <- as.data.frame(as.table(train_conf$table))
names(conf_matrix_data) <- c("Predicted", "Actual", "Freq")

# calculate relevant statistics
conf_matrix_data <- conf_matrix_data %>%
  group_by(Actual) %>%
  mutate(
    Percentage = Freq / sum(Freq) * 100,
    Label = paste0(Freq, "\n(", round(Percentage, 1), "%)")
  )

# plot heatmap ofr confusion matrix
ggplot(conf_matrix_data, aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile(color = "white", alpha = 0.8) +
  geom_text(aes(label = Label), color = "black", size = 5) +
  scale_fill_gradient(low = "#F8F9FA", high = "blue", name = "frequency") +
  labs(
    title = "heatmap ofr confusion matrix",
    subtitle = paste("Accuracy:", round(train_conf$overall["Accuracy"]*100, 1), "%"),
    x = "Actural class",
    y = "Predicted class",
    caption = paste("sensitivity:", round(train_conf$byClass["Sensitivity"]*100, 1), "% |",
                   "Specificity:", round(train_conf$byClass["Specificity"]*100, 1), "%")
  ) +
  theme_minimal() 
```

## **6. Answering data science questions**

#### How to identify potential mental issue, In other words, what features are most predictive of needing mental health treatment?

-   Employees with family history, reporting frequent work interference, and lack of mental care are more likely to need treatment.

#### What could employers do reducing risks?

-   A logistic regression model can reasonably predict employees at higher risk and support HR in targeted interventions.

<!-- -->

-   Increasing mental health care benefits, easy-access mental care treatment resources,

-   Companies should focus on improving support structures for employees showing risk factors.

```{r}
# Simplified risk scoring suggestion for workplace HR
risk_factors <- feature_importance %>%
  select(term, odds_ratio, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(desc(abs(log(odds_ratio))))

#Key variables increasing risk of needing treatment
risk_factors
```

# **7.**

-   We identified that **family history**, **work interference**, and **workplace support systems** are among the strongest predictors of mental health treatment need.
-   **Visualizations** in this report highlight key risk factors and their relationship with treatment-seeking.
-   **Limitations:** Subjective survey responses, possible response bias, and missingness may affect generalizability.
-   Future work could include more advanced ML (random forest, boosting), longitudinal data, and testing interventions.

------------------------------------------------------------------------

> **References** - OSMI Mental Health in Tech Survey [Kaggle Dataset](https://www.kaggle.com/datasets/osmi/mental-health-in-tech-survey/data) - [R for Data Science](https://r4ds.had.co.nz/) - [Tidyverse Documentation](https://www.tidyverse.org/) - [Introduction to Statistical Learning](https://www.statlearning.com/)

# Simple correlation to interaction effects

```{r}
# 1. Interaction Effects Visualization - work_interfere * family_history
interaction_data <- train_data %>%
  group_by(work_interfere, family_history) %>%
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  ) %>%
  filter(!is.na(work_interfere))

ggplot(interaction_data, aes(x = work_interfere, y = treatment_rate, 
                             color = family_history, group = family_history)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            vjust = -0.8, size = 3) +
  labs(
    title = "Interaction Effect: Work Interference × Family History",
    subtitle = "Family history significantly amplifies the impact of work interference on treatment needs",
    x = "Frequency of Work Interference",
    y = "Treatment Rate",
    color = "Family History of\nMental Health"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("No" = "#E74C3C", "Yes" = "#3498DB"))
```

```{r}
# 2. Interaction Effects Visualization - benefits * Gender
gender_benefits_data <- train_data %>%
  group_by(Gender, benefits) %>%
  summarise(
    treatment_rate = mean(treatment_binary),
    n = n(),
    .groups = 'drop'
  )

ggplot(gender_benefits_data, aes(x = Gender, y = treatment_rate, 
                                fill = benefits)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(round(treatment_rate*100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(
    title = "Interaction Effect: Gender × Benefits Access",
    subtitle = "Male treatment rates drop significantly without benefits, while females show more resilience",
    x = "Gender",
    y = "Treatment Rate",
    fill = "Company Provides\nMental Health Benefits"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set2")
```

**"When we dug deeper into our analysis, we discovered a crucial insight: risk factors don't operate in isolation - they amplify or buffer each other.”**

**The first chart reveals** a powerful interaction between **work interference and family history**:

-   For employees without family history, even frequent work interference leads to only 60% treatment rate

-   But for those with family history, the same level of work interference drives treatment rates above 90%

-   **Business Insight**: This tells us screening should **prioritize employees who report both family history AND work interference**

**The second chart shows** how **gender and (mental health)benefits** interact to reveal important organizational patterns:

-   Male employees are highly sensitive to benefits availability - treatment rates drop sharply without benefits

```{r}
# 3. Non-linear Relationships Visualization - Age
library(mgcv)
install.packages("gratia")
library(gratia)

# Fit GAM model
gam_age <- gam(treatment_binary ~ s(Age) + work_interfere + family_history,
               data = train_data, family = "binomial", method = "REML")

# Create prediction data
age_range <- seq(min(train_data$Age, na.rm = TRUE), 
                max(train_data$Age, na.rm = TRUE), length.out = 100)
pred_data <- data.frame(
  Age = age_range,
  work_interfere = factor("Sometimes", levels = levels(train_data$work_interfere)),
  family_history = factor("No", levels = levels(train_data$family_history))
)

# Get predictions
predictions <- predict(gam_age, newdata = pred_data, type = "link", se.fit = TRUE)
pred_data$fit <- plogis(predictions$fit)
pred_data$upper <- plogis(predictions$fit + 1.96 * predictions$se.fit)
pred_data$lower <- plogis(predictions$fit - 1.96 * predictions$se.fit)

# Plot non-linear relationship
ggplot(pred_data, aes(x = Age, y = fit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#3498DB") +
  geom_line(size = 1.5, color = "#2980B9") +
  labs(
    title = "Non-linear Relationship: Age vs Treatment Needs",
    subtitle = "Risk peaks in early adulthood and mid-life, challenging simple linear assumptions",
    x = "Age",
    y = "Predicted Treatment Probability"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent)
```

```{r}
# 4. Create comprehensive dashboard of all advanced analyses
library(patchwork)

# Prepare four charts
p1 <- ggplot(interaction_data, aes(x = work_interfere, y = treatment_rate, 
                                  color = family_history, group = family_history)) +
  geom_line(size = 1) + geom_point(size = 2) +
  labs(title = "Interaction 1: Work Interference × Family History", x = NULL, y = "Treatment Rate") +
  theme_minimal() + scale_y_continuous(labels = scales::percent)

p2 <- ggplot(gender_benefits_data, aes(x = Gender, y = treatment_rate, fill = benefits)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Interaction 2: Gender × Benefits Access", x = NULL, y = "Treatment Rate") +
  theme_minimal() + scale_y_continuous(labels = scales::percent)

p3 <- ggplot(pred_data, aes(x = Age, y = fit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  geom_line(size = 1, color = "#2980B9") +
  labs(title = "Non-linear Effect: Age Pattern", x = "Age", y = "Predicted Probability") +
  theme_minimal()

# Random Forest feature importance
rf_importance <- as.data.frame(rf_model$importance)
rf_importance$Variable <- rownames(rf_importance)
rf_importance <- rf_importance %>% arrange(desc(MeanDecreaseGini))

p4 <- ggplot(rf_importance[1:6, ], aes(x = reorder(Variable, MeanDecreaseGini), 
                                      y = MeanDecreaseGini)) +
  geom_col(fill = "#27AE60") +
  coord_flip() +
  labs(title = "Multi-method Validation: Random Forest Feature Importance", 
       x = NULL, y = "Importance Score (Gini Decrease)") +
  theme_minimal()

# Combine charts
(p1 + p2) / (p3 + p4) +
  plot_annotation(
    title = "Advanced Analytics: Four Key Insights Beyond Simple Correlations",
    theme = theme(plot.title = element_text(size = 16, face = "bold"))
  )
```

```{r correlation-heatmap, warning=FALSE, message=FALSE}
# 说明：
# - 我们把因子变量临时编码为数值（as.numeric(as.factor(.))）来计算 Spearman 相关系数。
# - 这只是用于“可视化与探索”目的：因子->数值的转换会改变变量含义，不能当作严格的相关性/因果性证据。
# - 对于类别-类别的关联，更合适的方法是 Cramer's V；对于数值-二分类可用 point-biserial 等。
#
# 需要的包：corrplot, reshape2, Hmisc（可选，用于 p 值）
if(!requireNamespace("corrplot", quietly=TRUE)) install.packages("corrplot")
if(!requireNamespace("reshape2", quietly=TRUE)) install.packages("reshape2")
if(!requireNamespace("Hmisc", quietly=TRUE)) install.packages("Hmisc")

library(corrplot); library(reshape2); library(Hmisc)

# 选取用于可视化的变量（使用在建模中挑选的 predictors）
corr_df <- model_data 

# 把因子变量临时编码为数值（仅用于计算相关矩阵）
corr_num <- corr_df |>
  mutate(across(everything(), ~ if(is.factor(.)) as.numeric(as.factor(.)) else .))

# 计算 Spearman 相关系数（对有序/非正态数据更稳健）
corr_mat <- cor(corr_num, use = "pairwise.complete.obs", method = "spearman")

corr_mat[upper.tri(corr_mat_masked, diag = TRUE)] <- NA

# 备用：用 ggplot2 绘制完整的热图（旋转标签更友好）
melted <- reshape2::melt(corr_mat, na.rm = TRUE)
ggplot(melted, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limits = c(-0.5,0.6)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "", fill = "Spearman\nrho",
       title = "Spearman Correlation Heatmap (factors -> numeric)")
```
